A SIMPLE AND EFFICIENT ALGORITHM FOR COMPUTING APPROXIMATE NASH-EQUAILIBRIA IN SEQUENTIAL GAMES WITH INCOMPLETE INFORMATION

### 2015-09-26 paper.tex
added ieee copyright contract

% Title.
% ------
\title{ A simple and efficient algorithm for computing approximate
  Nash equilibria in sequential games with incomplete
  information}

@@

\maketitle
\begin{abstract}We present a simple primal-dual algorithm for
  computing approximate Nash equilibria in two-person zero-sum
  sequential games with  incomplete information and perfect recall
  (like Texas Hold'em poker). Our algorithm only performs basic
  iterations (i.e matvec multiplications, clipping, etc., and no calls
  to external first-order oracles, no matrix inversions, etc.)
  and is applicable to a broad class of two-person zero-sum games
  including simultaneous games and sequential games with incomplete
  information and perfect recall. The applicability to the latter kind
  of games is thanks to the sequence-form representation
  \cite{koller1992complexity} which allows us to encode any such game
  as a matrix game with convex polytopial profiles. We prove that
  the number of iterations needed to produce a Nash equilibrium with a
  given precision is inversely proportional to the
  precision. We also present experimental results on simulated and
  real games (Kuhn poker).
\end{abstract}

\textbf{keywords}: algorithmic-game theory; sequential game;
incomplete information; perfect recall; approximate Nash equilibrium;
primal-dual algorithm; convex-optimization

@@

The sequence-form representation for two-person zero-sum games with
incomplete information was introduced in
\cite{koller1992complexity}, and the theory was further developed in
\cite{koller1994fast,von1996efficient,vonequilibrium} where it was
established that for such games, there exist sparse matrices
$A \in \mathbb{R}^{n_1 \times n_2}$, $E_1 \in \mathbb{R}^{l_1 \times
  n_1}$, $E_2 \in \mathbb{R}^{l_2 \times n_2}$, and vectors $e_1 \in
\mathbb{R}^{l_1}, e_2 \in \mathbb{R}^{l_2}$ such that $n_1$, $n_2$,
$l_1$, and $l_2$ are all linear in the size of the game tree (number
of states in the game) and such that Nash equilibria correspond to
pairs $(x, y)$ of \textit{realization plans} which solve the primal
LCP (Linear Convex Program)
\begin{equation}
  \begin{aligned}
    \underset{(y,p) \in \mathbb{R}^{n_2} \times
     \mathbb{R}^{l_1}}{\text{minimize }}\langle e_1,
    p\rangle \hspace{.5em}\text{
       subject to: } &y \ge 0, E_2y = e_2,\\
    &-Ay + E_1^Tp \geq 0,
  \end{aligned}
  \label{eq:primal_pb}
\end{equation}

@@

Solving the LCPs \eqref{eq:primal_pb} and \eqref{eq:dual_pb} exactly
is impossible in practice (indeed, this system of problems is NP-hard
\cite{koller1992complexity}) and such a precision doesn't have any
fundamental practical advantage. Instead, it is customary compute
``approximate'' Nash equilibria. A popular notion of approximate
equilibria is the following:

@@

Developing on an alternative notion of approximate equilibria (see
Definition \ref{thm:cool_notion})
homologous to that presented in Definition \ref{thm:approx_nash}, we
device a primal-dual algorithm that (Algorithm \ref{Tab:algo}) for
computing approximate Nash equilibria in sequential two-person
zero-sum games with incomplete information and perfect recall, and
which satisfies the above crucial requirements (1)--(2). We also
prove (Theorem \ref{thm:pd}) that --in an ergodic / Ces\`ario sense-- the number of
iterations required by the algorithm to produce an approximation
equilibrium to a precision $\epsilon$ is $\mathcal{O}(1/\epsilon)$,
with explicit values for the constants involved this worst-case cost.
These contributions will be elaborated in section \ref{sec:gsp}.
%%  The source of approximation in our scheme
%% is that we explicitly control how close we are to the feasible set
%% $Q_1 \times Q_2$.

@@

Then GSP($\Psi_1$,
  $\Psi_2$, $g_1$, $g_2$) is equivalent to the Nash equilibrium LCPs
  \eqref{eq:primal_pb} and \eqref{eq:dual_pb}, i.e
a quadruplet $(y^*,p^*, x^*, q^*) \in \mathbb{R}^{n_2}
  \times \mathbb{R}^{l_1} \times \mathbb{R}^{n_1} \times
  \mathbb{R}^{l_2}$ solves the Nash equilibrium LCPs
  \eqref{eq:primal_pb} and \eqref{eq:dual_pb} iff it solves
  GSP($\Psi_1$, $\Psi_2$, $g_1$, $g_2$). 
  \label{thm:pd}
\end{theorem}

@@

By Theorem \ref{thm:pd}, solving for a Nash equilibrium for
the game is equivalent to solving the GSP
\eqref{eq:unconstrained_pb}, which as it turns out, is simpler
conceptually (for example, we no longer need to compute the
complicated orthogonal projections $\proj_{Q_k}$). The rest of the
paper will be devoted to developing an algorithm for solving the latter.

@@

\subsection{A more comfortable notion of approximate Nash equilibrium}
\begin{definition}
  Given a scalar $\epsilon > 0$ and a function $f:\mathbb{R}^n
  \rightarrow [-\infty,+\infty]$, the $\epsilon$-enlarged
  subdifferential (or
  $\epsilon$-subdifferential, for short) of $f$ is the set-valued
  function defined by
  \begin{eqnarray}
\partial_\epsilon f(x):= \{s \in \mathbb{R}^n | f(z)
\ge f(x) + \langle s, z - x\rangle - \epsilon,\forall z \in
\mathbb{R}^n\}.
\end{eqnarray}
\end{definition}
The idea behind $\epsilon$-subdifferentials is the following. Say we wish
to minimize the function $f$. Replace the usual necessary
condition ``$0 \in \partial f(x)$'' for the optimality of $x$ with the
weaker condition ``$\partial_\epsilon f(x)$ contains a sufficiently
small vector $v$''. In fact, it is easy to see that, for each
point $x \in \mathbb{R}^n$, we have $\underset{\epsilon \rightarrow
  0^+}{\text{lim }}\partial_\epsilon f(x) = \partial f(x)$.

This approximation concept for subdifferentials yields the following
concept of approximate Nash equlibria (adapted from
\cite{he2013accelerating}).

@@

The above definition is a generalization of the notion of Nash
equilibria since:
\begin{itemize}
\item Exact Nash equilibria correspond to Nash $(0,0)$-equilibria.
\item Nash $\epsilon$-equilibria (in the sense of Definition
  \ref{thm:approx_nash}) correspond to Nash $(0,\epsilon)$-equilibria.
\end{itemize}

@@

\section{Concluding remarks and future work}
Making use of the sequence-form representation
\cite{koller1992complexity,von1996efficient,vonequilibrium}, we have
devised a simple and efficient primal-dual algorithm for computing
Nash equilibria in two-person zero-sum sequential games with
incomplete information (like Texas Hold'em, etc.). Our algorithm is
simple to implement, with a low constant cost per iteration, and
enjoys a rigorous convergence theory with a proven
$\mathcal{O}(1/\epsilon)$ convergence in terms of basic operations
(matvec products, clipping, etc.), to a Nash
$(\epsilon,0)$-equilibrium of the game.

>>>>

% Title.
% ------
\title{A SIMPLE AND EFFICIENT ALGORITHM FOR COMPUTING APPROXIMATE
  NASH-EQUAILIBRIA IN SEQUENTIAL GAMES WITH INCOMPLETE INFORMATION}

@@

\maketitle
\begin{abstract}
We present a simple primal-dual algorithm for
  computing approximate Nash-equilibria in two-person zero-sum
  sequential games with  incomplete information and perfect recall
  (like Texas Hold'em poker). Our algorithm only performs basic
  iterations (i.e matvec multiplications, clipping, etc., and no calls
  to external first-order oracles, no matrix inversions, etc.)
  and is applicable to a broad class of two-person zero-sum games
  including simultaneous games and sequential games with incomplete
  information and perfect recall. The applicability to the latter kind
  of games is thanks to the sequence-form representation
  \cite{koller1992complexity} which allows us to encode any such game
  as a matrix game with convex polytopial profiles. We prove that
  the number of iterations needed to produce a Nash-equilibrium with a
  given precision is inversely proportional to the
  precision. We also present experimental results on simulated and
  real games (Kuhn poker).
\end{abstract}

\begin{keywords} algorithmic-game theory, sequential games,
incomplete information, perfect recall, approximate Nash-equilibrium
\end{keywords}

@@

The sequence-form representation for two-person zero-sum games with
incomplete information was introduced in
\cite{koller1992complexity}, and the theory was further developed in
\cite{koller1994fast,von1996efficient,vonequilibrium} where it was
established that for such games, there exist sparse matrices
$A \in \mathbb{R}^{n_1 \times n_2}$, $E_1 \in \mathbb{R}^{l_1 \times
  n_1}$, $E_2 \in \mathbb{R}^{l_2 \times n_2}$, and vectors $e_1 \in
\mathbb{R}^{l_1}, e_2 \in \mathbb{R}^{l_2}$ such that $n_1$, $n_2$,
$l_1$, and $l_2$ are all linear in the size of the game tree (number
of states in the game) and such that Nash-equilibria correspond to
pairs $(x, y)$ of \textit{realization plans} which solve the primal
LCP (Linear Convex Program)
\begin{equation}
  \begin{aligned}
    \underset{(y,p) \in \mathbb{R}^{n_2} \times
     \mathbb{R}^{l_1}}{\text{minimize }}\langle e_1,
    p\rangle \hspace{.5em}\text{
       subject to: } &y \ge 0, E_2y = e_2,\\
    &-Ay + E_1^Tp \geq 0,
  \end{aligned}
  \label{eq:primal_pb}
\end{equation}

@@

Solving the LCPs \eqref{eq:primal_pb} and \eqref{eq:dual_pb} exactly
is impossible in practice (indeed, this system of problems is NP-hard
\cite{koller1992complexity}) and such a precision doesn't have any
fundamental practical advantage. Instead, it is customary compute
``approximate'' Nash-equilibria. A popular notion of approximate
equilibria is the following:

@@

Developing on an alternative notion of approximate equilibria (see
Definition \ref{thm:cool_notion})
homologous to that presented in Definition \ref{thm:approx_nash}, we
device a primal-dual algorithm that (Algorithm \ref{Tab:algo}) for
computing approximate Nash-equilibria in sequential two-person
zero-sum games with incomplete information and perfect recall, and
which satisfies the above crucial requirements (1)--(2). We also
prove (Theorem \ref{thm:pd}) that --in an ergodic / Ces\`ario sense-- the number of
iterations required by the algorithm to produce an approximation
equilibrium to a precision $\epsilon$ is $\mathcal{O}(1/\epsilon)$,
with explicit values for the constants involved this worst-case cost.
These contributions will be elaborated in section \ref{sec:gsp}.
%%  The source of approximation in our scheme
%% is that we explicitly control how close we are to the feasible set
%% $Q_1 \times Q_2$.

@@

Then GSP($\Psi_1$,
  $\Psi_2$, $g_1$, $g_2$) is equivalent to the Nash-equilibrium LCPs
  \eqref{eq:primal_pb} and \eqref{eq:dual_pb}, i.e
a quadruplet $(y^*,p^*, x^*, q^*) \in \mathbb{R}^{n_2}
  \times \mathbb{R}^{l_1} \times \mathbb{R}^{n_1} \times
  \mathbb{R}^{l_2}$ solves the Nash-equilibrium LCPs
  \eqref{eq:primal_pb} and \eqref{eq:dual_pb} iff it solves
  GSP($\Psi_1$, $\Psi_2$, $g_1$, $g_2$). 
  \label{thm:pd}
\end{theorem}

@@

By Theorem \ref{thm:pd}, solving for a Nash-equilibrium for
the game is equivalent to solving the GSP
\eqref{eq:unconstrained_pb}, which as it turns out, is simpler
conceptually (for example, we no longer need to compute the
complicated orthogonal projections $\proj_{Q_k}$). The rest of the
paper will be devoted to developing an algorithm for solving the latter.

@@

\subsection{A more comfortable notion of approximate Nash-equilibrium}
\begin{definition}
  Given a scalar $\epsilon > 0$ and a function $f:\mathbb{R}^n
  \rightarrow [-\infty,+\infty]$, the $\epsilon$-enlarged
  subdifferential (or
  $\epsilon$-subdifferential, for short) of $f$ is the set-valued
  function defined by
  \begin{eqnarray}
\partial_\epsilon f(x):= \{s \in \mathbb{R}^n | f(z)
\ge f(x) + \langle s, z - x\rangle - \epsilon,\forall z \in
\mathbb{R}^n\}.
\end{eqnarray}
\end{definition}
The idea behind $\epsilon$-subdifferentials is the following. Say we wish
to minimize the function $f$. Replace the usual necessary
condition ``$0 \in \partial f(x)$'' for the optimality of $x$ with the
weaker condition ``$\partial_\epsilon f(x)$ contains a sufficiently
small vector $v$''. In fact, it is easy to see that, for each
point $x \in \mathbb{R}^n$, we have $\underset{\epsilon \rightarrow
  0^+}{\text{lim }}\partial_\epsilon f(x) = \partial f(x)$.

This approximation concept for subdifferentials yields the following
concept of approximate Nash-equlibria (adapted from
\cite{he2013accelerating}).

@@

The above definition is a generalization of the notion of Nash
equilibria since:
\begin{itemize}
\item Exact Nash-equilibria correspond to Nash $(0,0)$-equilibria.
\item Nash $\epsilon$-equilibria (in the sense of Definition
  \ref{thm:approx_nash}) correspond to Nash $(0,\epsilon)$-equilibria.
\end{itemize}

@@

\section{Concluding remarks and future work}
Making use of the sequence-form representation
\cite{koller1992complexity,von1996efficient,vonequilibrium}, we have
devised a simple and efficient primal-dual algorithm for computing
Nash-equilibria in two-person zero-sum sequential games with
incomplete information (like Texas Hold'em, etc.). Our algorithm is
simple to implement, with a low constant cost per iteration, and
enjoys a rigorous convergence theory with a proven
$\mathcal{O}(1/\epsilon)$ convergence in terms of basic operations
(matvec products, clipping, etc.), to a Nash
$(\epsilon,0)$-equilibrium of the game.
